---
title: "STA 445 Assignment Five"
author: "Calder Evans"
date: "9 November 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(viridis)
library(latex2exp)
library(plotly)
library(ggrepel)
```

## Chapter Fourteen
### Exercise One
The `infmort` data set from the package `faraway` gives the infant mortality
rate for a variety of countries. The information is relatively out of date
(from 1970s?), but will be fun to graph. Visualize the data using by creating
scatter plots of mortality vs income while faceting using `region` and setting
color by `oil` export status. Utilize a $\log_{10}$ transformation for both
`mortality` and `income` axes. This can be done either by doing the
transformation inside the `aes()` command or by utilizing the `scale_x_log10()`
or `scale_y_log10()` layers. The critical difference is if the scales are on
the original vs log transformed scale. Experiment with both and see which you
prefer.
    
  a)  The `rownames()` of the table gives the country names and you should
        create a new column that contains the country names. *`rownames`
```{r}
    Infant.Mortality <- faraway::infmort %>% 
      mutate(country = rownames(faraway::infmort))
```

\newpage
  b)  Create scatter plots with the `log10()` transformation inside the `aes()`
        command.
```{r}
      ggplot(Infant.Mortality, aes(x=log10(income), y=log10(mortality))) +
      geom_point(aes(color = oil)) +
      facet_wrap(~region) + 
      labs(title="Infant Mortality vs. Income", x="Income", y="Infant Mortality")
```

\newpage
  c)  Create the scatter plots using the `scale_x_log10()` and `scale_y_log10()`.
      Set the major and minor breaks to be useful and aesthetically pleasing.
      Comment on which version you find easier to read.
```{r}
      ggplot(Infant.Mortality, aes(x = income, y = mortality)) +
      scale_x_log10() + 
      scale_y_log10() +
      geom_point(aes(color = oil)) +
      facet_wrap(~region) + 
      labs(title="Infant Mortality vs. Income", x="Income", y="Infant Mortality")
```

\newpage
  d)  The package `ggrepel` contains functions `geom_text_repel()` and
      `geom_label_repel()` that mimic the basic `geom_text()` and `geom_label()`
      functions in `ggplot2`, but work to make sure the labels don't overlap.
      Select 10-15 countries to label and do so using the `geom_text_repel()`
      function.
```{r}
    ggplot(Infant.Mortality, aes(x = income, y = mortality)) +
    scale_x_log10() + 
    scale_y_log10() +
    geom_point(aes(color = oil)) +
    facet_wrap(~region) + 
    labs(title="Infant Mortality vs. Income", x="Income", y="Infant Mortality") +
    geom_text_repel(aes(label=country))
```


\newpage
### Exercise Two
Using the `datasets::trees` data, complete the following:
  a)  Create a regression model for $y=$ `Volume` as a function of $x=$ `Height`.
```{r}
    model <- lm( Volume ~ Height, data=trees)
    trees <- trees %>% mutate(fit=fitted(model))
```

  b)  Using the `summary` command, get the y-intercept and slope of the
      regression line.
```{r}
    summary(model)
```

\newpage
  c)  Using `ggplot2`, create a scatter plot of Volume vs Height. 
```{r}
    ggplot( data=trees, aes(x = Height, y = Volume)) + geom_point()
```

\newpage
  d)  Create a nice white filled rectangle to add text information to using by
      adding the following annotation layer.

```{r}
    ggplot( data=trees, aes(x = Height, y = Volume)) + geom_point() +
    annotate('rect', xmin=65, xmax=75, ymin=57.5, ymax=72.5,
        fill='white', color='black')
```

\newpage
  e)  Add some annotation text to write the equation of the line
      $\hat{y}_i = -87.12 + 1.54 * x_i$ in the text area.
```{r}
    ggplot( data=trees, aes(x = Height, y = Volume)) + geom_point() +
    annotate('rect', xmin=65, xmax=75, ymin=57.5, ymax=72.5,
        fill='white', color='black') +
    annotate("text", x=70, y=68, 
             label=latex2exp::TeX("$\\hat{y}_i = -87.12 + 1.54x_{i}$"))    
```

\newpage
  f)  Add annotation to add $R^2 = 0.358$
```{r}
    ggplot( data=trees, aes(x = Height, y = Volume)) + geom_point() +
    annotate('rect', xmin=65, xmax=75, ymin=57.5, ymax=72.5,
             fill='white', color='black') +
    annotate("text", x=70, y=68, 
             label=latex2exp::TeX("$\\hat{y}_i = -87.12 + 1.54x_{i}$")) + 
    annotate("text", x=70, y=64,
             label=latex2exp::TeX("$R^2 = 0.358$"))
```

\newpage
  g)  Add the regression line in red. The most convenient layer function to uses
        is `geom_abline()`. It appears that the `annotate` doesn't work with
        `geom_abline()` so you'll have to call it directly.
```{r}
    ggplot( data=trees, aes(x = Height, y = Volume)) + geom_point() +
    annotate('rect', xmin=65, xmax=75, ymin=57.5, ymax=72.5,
             fill='white', color='black') +
    annotate("text", x=70, y=68, 
             label=latex2exp::TeX("$\\hat{y}_i = -87.12 + 1.54x_{i}$")) + 
    annotate("text", x=70, y=64,
             label=latex2exp::TeX("$R^2 = 0.358$")) +
    geom_line(aes(y=fit, color="red")) +
    theme(legend.position= "none")
```


