---
title: "445 Assignment One"
author: "Calder Evans"
date: "6 October 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Chapter Eight

### Exercise One 
Create a vector of three elements (2,4,6) and name that vector `vec_a`. 
Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors 
together and name the result `vec_c`.

```{r}
    vec_a <- c( 2, 4, 6)
    vec_b <- c( 8, 10, 12)
    vec_c <- vec_a + vec_b
      vec_c
```

### Exercise Two
Create a vector, named `vec_d`, that contains only two elements (14,20). 
Add this vector to `vec_a`. What is the result and what do you think R did 
(look up the recycling rule using Google)? What is the warning message that 
R gives you?

```{r}
    vec_d <- c(14,20)
    vec_d + vec_a
```

  (a) The result is a vector with three numbers. The first and second values
      are computed relatively easily. The first value in the printed vector
      results from adding the first value of vector d with the first value of 
      vector a . The second value in the printed vector results from adding the 
      second value of vector d with the second value of vector a. The third 
      number "recycles" the first value of vector d and adds it to the third 
      value of vector a.
  (b) The warning message is reminding you that the vectors are not the same 
      size.

### Exercise Three
Next add 5 to the vector vec_a. What is the result and what did R do? Why 
doesn't it give you a warning message similar to what you saw in the previous 
problem?

```{r}
    5 + vec_a
```

  a) The result adds 5 to every value in the a vector
  b) This sum does not give you a error like the previous problem because it is
     computing vector algebra, where you add 5 to every value.

### Exercise Four
Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two different 
ways. 
    a) First using the `seq()` function 
    b) Using the `a:b` shortcut.

```{r}
    seq( from=1, to=5, by=1)
    1:5
```

### Exercise Five
Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 
    a) Using the seq() function and 
    b) Using the a:b shortcut and some subsequent algebra. 
       *Hint: Generate the vector 1-10 and then multiple it by 2*.

```{r}
    seq( from=2, to=20, by=2)
    2 * 1:10
```

### Exercise Six
Generate a vector of 21 elements that are evenly placed between 0 and 1 using 
the `seq()` command and name this vector `x`. 

```{r}
    x <- seq( from=0, to=1, by=0.05)
      x
```

### Exercise Seven
Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ using the `rep()` 
command to replicate the vector c(2,4,8). 

```{r}
    rep( c(2,4,8), 3)
```

### Exercise Eight
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$ using the 
`rep()` command. You might need to check the help file for rep() to see all of 
the options that rep() will accept. In particular, look at the optional argument 
`each=`.

```{r}
    rep( c(2,4,8), each=3)
```
\newpage
### Exercise Ten
a) Create the matrix in two ways and save the resulting matrix as `M`.
    i. Create the matrix using some combination of the `seq()` and 
       `matrix()` commands.
       
```{r}
    M <- matrix(seq( from=2, to=30, by=2), nrow=3, ncol=5)
      M
```

    ii. Create the same matrix by some combination of multiple `seq()` 
        commands and either the `rbind()` or `cbind()` command.

```{r}
    a <- seq( from=2, to=10, by=2)
    b <- seq( from=12, to=20, by=2)
    c <- seq( from=22, to=30, by=2)
    M <- rbind( a, b, c)
      M
```

b) Extract the second row out of `M`.

```{r}
    M[2,]
```
c) Extract the element in the third row and second column of `M`.

```{r}
    M[3,2]
```

### Exercise Twelve
The following code creates a `data.frame` and then has two different methods 
for removing the rows with `NA` values in the column `Grade`. 
Explain the difference between the two.
```{r, eval=FALSE}
    df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
    df[ -which(  is.na(df$Grade) ), ]
    df[  which( !is.na(df$Grade) ), ]
```

(a) The subtraction sign in the first method subtracts the "NA" values from the 
    data set.
    The exclamation point in the second method is a "not" operator and it takes
    out all the values that aren't "NA"
\newpage

### Exercise Fourteen
Create and manipulate a list.
  a) Create a list named my.test with elements
      + x = c(4,5,6,7,8,9,10)
      + y = c(34,35,41,40,45,47,51)
      + slope = 2.82
      + p.value = 0.000131
```{r}
    mytest <- list( x=c(4,5,6,7,8,9,10), 
                    y=c(34,35,41,40,45,47,51), 
                    slope=2.82, 
                    p.value=0.000131)
      mytest
```

    b) Extract the second element in the list.
```{r}
    mytest[[2]]
```

    c) Extract the element named `p.value` from the list.
```{r}
    mytest$p.value
```
\newpage

## Chapter Nine
### Exercise One
Download from GitHub the data file [Example_5.xls]
(https://github.com/dereksonderegger/444/raw/master/data-raw/Example_5.xls). 
Open it in Excel and figure out which sheet of data we should import into R. 
At the same time figure out how many initial rows need to be skipped. Import 
the data set into a data frame and show the structure of the imported data 
using the `str()` command. Make sure that your data has $n=31$ observations
and the three columns are appropriately named. If you make any modifications 
to the data file, comment on those modifications.


```{r}
    library(readxl)
    data.trees <- data.frame(read_excel('Example_5.xls', 
                                        sheet='RawData', 
                                        range=("A5:C36")))
      head(data.trees)
```

### Exercise Two
Download from GitHub the data file [Example_3.xls]
(https://github.com/dereksonderegger/444/raw/master/data-raw/Example_3.xls). 
Import the data set into a data frame and show the structure of the imported 
data using the `tail()` command which shows the last few rows of a data table. 
Make sure the Tesla values are `NA` where appropriate and that both 
`-9999` and `NA` are imported as NA values. If you make any modifications to 
the data file, comment on those modifications.

```{r}
    data.cars <- read_excel('Example_3.xls', sheet=2, range=("A1:L34"), na=c("NA", "-9999"))
      tail(data.cars)
```

